From Big Data to Big Displays\\ High-Performance Visualization at Blue Brain

### 2017-06-02 paper.tex
RTNeuron description and review up to that point.

  We present the key elements of this HPV ecosystem, which integrates C++
  visualization applications with novel collaborative display systems. We
  motivate how our strategy of transforming visualization engines into services
  enables a variety of use cases, not only for the integration with
  high-fidelity displays, but also to build service oriented architectures, to
  link into web applications and to provide remote services to python
  applications.
\end{abstract}

@@

Visualization supports the BBP along all parts of the project to understand and
debug model data, building and simulation algorithms as well as validating and
discovering new insight from the in-silico experiments. Our strategy to support
this mission is based on components linked through network protocols:
High-fidelity display systems to see more detail in complex data, a set of
standard rendering engines (rasterization, out-of-core volume rendering,
interactive raytracing), and decoupled, light-weight applications using remote
these components. In the following we will present these components along with a
few use cases.

@@

High-fidelity display systems are the integration point of the Blue Brain
visualization capabilities. They are the evolution of existing visualization
systems, enabling higher resolution, immersion and team collaboration. Compared
to current single-user or single-presenter systems, collaborative display
systems allow real team work through a combination of size, resolution and user
friendly implementation. Compared to immersive visualization systems like the
CAVE, they provide a more approachable environment for high-fidelity
visualization. For all use cases, the increased display size and resolution
allows better data exploration for 2D and 3D content, facilitating large data
exploration.

Tide (Tiled Interactive Display Environment) is the software driving the Blue
Brain tiled display walls and the future OpenDeck. It provides multi-window,
multi-user touch interaction on large surfaces --- think of a giant
collaborative wall-mounted tablet. Tide is a distributed application that can
run on multiple machines to power display walls or projection systems of any
size. Its user interface is designed to offer an intuitive experience on touch
walls. It works just as well on non touch-capable installations by using its web
interface from any web browser. \fig{fTDW} shows Tide on a $4\times 3$ display
wall and \fig{fTideWeb} shows the Tide web interface in a browser.

While there is substantial research on tiled display wall software~\cite{Sage,
Sage2, Omegalib, DK:11, JLMV:06, DisplayCluster}, we found that most solutions
were not ready for production use in a $24\times 7$ unattended environment. For
this reason, we started with the TACC DisplayCluster open source
software~\cite{DisplayCluster}, which we incrementally refactored and improved
to the current TIDE implementation. On the other hand, the hardware has been
commoditized to make these type of installations affordable to medium-sized
institutions which allowed us to build the software integration for a reasonable
startup cost. We have focused on the multitouch user interface, which implements
a low entry barrier for new users, a unique capability of our solution.

@@

@@

Advances in computer hardware and software have brought raytracing to the point
where it replaces classical rasterization for virtually all use cases in
scientific visualization. On one hand, OpenGL-like local illumination for
typical data sets used with rasterization (up to hundred million triangles) can
be done at similar framerates to OpenGL~\cite{ospray}. For small datasets,
OpenGL performs better, but for larger data sets raytracing outperforms OpenGL.
This is due to a better scalability with respect to the data set size
($O(\log{N})$ vs $O(n)$). Furthermore, CPU-based raytracing allows the rendering
of larger data sets without any level-of-detail algorithms due to the larger
memory size. Last, but not least, advanced rendering algorithms such as
shadowing, reflections and global illuminations are significantly easier to
implement in a raytracer. The only area where rasterization provides a benefit
is for rendering at very high frame rates, needed for example for immersive
visualization.

It is for these reasons that interactive raytracing is the future technology for
interactive and offline rendering at Blue Brain. We have developed a first open
source implementation of a visualization engine with different backends:
OSPRay~\cite{ospray} for CPU-based raytracing and OptiX~\cite{optix} for
GPU-based raytracing. This application called Brayns can load and visualize our
data sets in a variety of modes and integrates with our messaging solution to be
accessible from Python. We are currently integrating the key algorithms from
Livre for out-of-core volume rendering in Brayns, which will also facilitate
mixing polygonal with volume data in a single scene.

@@

@@

%----------------------------------------------------------------------
\section*{Acknowledgments}
%----------------------------------------------------------------------
This publication was supported by the Blue Brain Project (BBP), the Swiss
National Science Foundation under Grant 200020-129525, the King Abdullah
University of Science and Technology (KAUST) through the KAUST-EPFL alliance for
Neuro-Inspired High Performance Computing, the Spanish Ministry of Science and
Innovation under grant (TIN2010-21289-C02-01/02), the Cajal Blue Brain Project,
Hasler Stiftung Projekt Nr. 12097, and from the European Union’s Horizon 2020
research and innovation programme under grant agreement No 720270 (HBP SGA1).

>>>>

  We present the key elements of this HPV ecosystem, which integrates C++
  visualization applications with novel collaborative display systems. We
  motivate how our strategy of transforming visualization engines into services
  enables a variety of use cases, not only for the integration with
  high-fidelity displays, but also to build service oriented architectures, to
  link into web applications and to provide remote services to Python
  applications.
\end{abstract}

@@

Visualization supports the BBP along all parts of the project to understand and
debug model data, building and simulation algorithms as well as validating and
discovering new insight from the in-silico experiments. Our strategy to support
this mission is based on components linked through network protocols:
High-fidelity display systems to see more detail in complex data, a set of
standard rendering engines (rasterization, out-of-core volume rendering,
interactive raytracing), and decoupled, light-weight applications using
these components remotely. In the following we will present these components
along with a few use cases.

@@

High-fidelity display systems are the integration point of the Blue Brain
visualization capabilities. They are the evolution of existing visualization
systems, enabling high resolution, immersion and team collaboration. Compared
to current single-user or single-presenter systems, collaborative display
systems allow real team work through a combination of size, resolution and user
friendly implementation. Compared to immersive visualization systems like the
CAVE, they provide a more approachable environment for high-fidelity
visualization. For all use cases, the increased display size and resolution
allows better data exploration for 2D and 3D content, facilitating large data
exploration.

@@

Tide (Tiled Interactive Display Environment) is the software driving the Blue
Brain tiled display walls. It provides multi-window, multi-user touch
interaction on large surfaces --- think of a giant collaborative wall-mounted
tablet. Tide is a distributed application that can run on multiple machines to
power display walls or projection systems of any size. Its user interface is
designed to offer an intuitive experience on touch walls. It works just as
well on non touch-capable installations by using its web interface from any
web browser. \fig{fTDW} shows Tide on a $4\times 3$ display wall and
\fig{fTideWeb} shows the Tide web interface in a browser.

While there is substantial research on tiled display wall software~\cite{Sage,
Sage2, Omegalib, DK:11, JLMV:06, DisplayCluster}, we found that most solutions
were not ready for production use in a $24\times 7$ unattended environment. For
this reason, we started with the TACC DisplayCluster open source
software~\cite{DisplayCluster}, which we incrementally refactored and improved
to the current TIDE implementation. On the other hand, the hardware has been
commoditized to make these type of installations affordable to medium-sized
institutions which allowed us to build the software integration for a reasonable
startup cost. % I don't understand very well what do you try to mean with this last sentence.
We have focused on the multitouch user interface, which implements
a low entry barrier for new users, a unique capability of our solution.

@@

Advances in computer hardware and software have brought raytracing to the point
where it replaces classical rasterization for virtually all use cases in
scientific visualization. On one hand, OpenGL-like local illumination for
typical data sets used with rasterization (up to hundred million triangles) can
be done at similar framerates to OpenGL~\cite{ospray}. For small datasets,
OpenGL performs better, but for larger data sets raytracing outperforms OpenGL.
This is due to a better scalability with respect to the data set size
($O(\log{N})$ vs $O(n)$). Furthermore, CPU-based raytracing allows the rendering
of larger data sets without any level-of-detail algorithms due to the larger
memory size. Last, but not least, advanced rendering algorithms such as
shadowing, reflections and global illuminations are significantly easier to
implement in a raytracer. The only area where rasterization provides a benefit
is for rendering at very high frame rates, needed for example for immersive
visualization, or very high resolutions.

It is for these reasons that interactive raytracing is the future technology for
interactive and offline rendering at Blue Brain. We have developed a first open
source implementation of a visualization engine with different backends:
OSPRay~\cite{ospray} for CPU-based raytracing and OptiX~\cite{optix} for
GPU-based raytracing. This application called Brayns can load and visualize our
data sets in a variety of modes and integrates with our messaging solution to be
accessible from Python. We are currently integrating the key algorithms from
Livre for out-of-core volume rendering in Brayns, which will also facilitate
mixing polygonal with large volumetric data in a single scene.

@@

RTNeuron is the oldest of our interactive visualization tools. Originally
conceived as a standalone application for visualizing simulation results, it
has evolved into a rasterization-based rendering engine library implemented in
C++ with a Python wrapping. The power of RTNeuron lies in a domain specific API
designed for the visualization of detailed neuronal circuit models, this API
allows for building custom applications tailored for specific use cases.
It provides features to visualize static circuit data and some simulation
results. Static data visualization includes different visual representations for
neurons, synapses, selective pruning of neuronal trees, clipping
planes... The simulation results that can be displayed are the spiking
activity of the neurons and scalar variables from the cable models mapped to
the neuron surface. The color maps for displaying these data are highly
configurable, being posible to apply different color maps to different cell
subsets.

Neurons can be displayed with different levels of detail ranging from simple
spheres, cylinder-like geometric models and detailed polygonal
meshes. Advanced rendering techniques included are several types of parallel
rendering algorithms (such as sort-first and sort-last) and several algorithms
for transparency rendering that enable efficient and correct rendering of
scenes with great geometrical complexity and high depth
complexity. Transparency is particularly suitable for masking and highlighting
features of interest on the circuit. Ongoing work is a more scalable parallel
rendering algorithm specially suited for transparency. Tiled rendering is also
possible, which allows very high resolution renderings suitable for printing
at sizes larger than A0.

The core engine is implemented in C++ using Equalizer and OpenSceneGraph and
also leverages part of our messaging framework to allow coupling with other
applications. The 3D view can be embedded inside Qt applications, in particular
in Python with PyQt and QML for overlaying GUI elements. Several use-case
specific applications have been built this way.

@@

%----------------------------------------------------------------------
\section*{Acknowledgments}
%----------------------------------------------------------------------

This publication was supported by the Blue Brain Project (BBP), the Swiss
National Science Foundation under Grant 200020-129525, the King Abdullah
University of Science and Technology (KAUST) through the KAUST-EPFL alliance for
Neuro-Inspired High Performance Computing, the Spanish Ministry of Science and
Innovation under grant (TIN2010-21289-C02-01/02), the Cajal Blue Brain Project,
Hasler Stiftung Projekt Nr. 12097, and from the European Union’s Horizon 2020
research and innovation programme under grant agreement No 720270 (HBP SGA1).

We would also like to thank the people from GMRV at the Rey Juan Carlos
University (URJC) for their collaboration under the Cajal Blue Brain and HBP
projects.
